Predicting Solid State Material Platforms for Quantum Technologies

### 2022-01-21 paper_twocolumn.tex
Updates from Overleaf

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[superscriptaddress,unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}

@@

We find that the Ferrenti approach and the extended Ferrenti approach predict a broad range of materials as suitable for quantum technologies. The restrictions set in the empirical approach, which relies exclusively on findings from the literature, gave substantially fewer candidates with a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. In contrast to the machine learning (ML) methods based on Ferrenti and extended Ferrenti approach that focused on band gap and ionic character, the ML methods from the empirical approach highlighted features related to symmetry and crystal structure, including bond length, orientation and radial distribution. All four machine learning (ML) methods applied to the training set derived using the empirical approach agreed on a set of $214$ predicted candidate materials.
All three approaches and all four machine learning methods agreed on a subset of $47$ eligible candidates %(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for further material explorations towards quantum technology.

@@

%Figure 1 introduction schematic 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/ht-workflow-new-2.png}
    \caption{Schematic of an example workflow in material informatics. Results from theory, experiment and computation are fed into material databases (arrow pointing to the center). A cycle involving material screening, machine learning and predictions leads to knowledge gain and ultimately applications in fields such as clean energy and quantum technology. 
    }
    \label{fig:ht-workflow}
\end{figure}

@@

Here we provide a framework for the data mining and automated discovery of promising solid-state semiconductor hosts for quantum emitters and spin centers using targeted database search and machine learning methods combined with knowledge from the field. \textit{En setning om discovery her??} 
In this framework we do not distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that high quality databases containing relevant theoretical and/or experimental data are available.

The approach we have developed relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
The Matminer Python library for data mining \cite{Ward2018} was then used for material analysis to featurize the extracted data. An important aspect of our work is the database building and pertinent development of the training sets for machine learning methods. We have developed three different approaches to data mining: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) 
%Marianne I have now changed augmented Ferrenti to extended Ferrenti 
\emph{the extended Ferrenti approach} and (iii) \emph{the empirical approach}. We find that the former two methodologies base their data extraction protocols on broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The third and empirical approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates.  The three training sets where then analyzed with the four supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting, and yielded $214$ predicted candidates that were common between all approaches and machine learning methods. Importantly, the approach emphasize features like the standard deviations  of  the  radial  distribution  function  (RDF) and average bond length as critical features for suitable QT compatible materials.
%(to a $0.5$ cut-off)  
Example materials include ZnGeP$_2$, MgSe, CdS, BP, BC$_2$N, BP, Ge, GeC, InP, and InAs. 
Focused theoretical and experimental studies, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, are needed to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart_2.png}
    \caption{The data flow of the main project, starting from an initial MP query, and ending with a featurized data set with entries from several other databases. To limit the memory and computational usage, the data are partitioned into smaller subsets where the respective Pymatgen objects  
    (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) 
    are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. Abbreviations used are Materials Project (MP), density of states (DOS) and electronic band structure (band).
    }
    \label{fig:flowchart}
\end{figure}

@@

\begin{figure}[t] %!htp
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \input{figures/parallel-coordinates-plots/01-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/02-augmented-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/03-insightful-approach-v2.pgf}
    \end{subfigure}
    %\vspace*{10mm}
    \caption{Parallel coordinate plots for the different approaches. To limit the data cluttering, we have randomly collected up to $250$ entries for each class and made the lines transparent. For the empirical approach, we have used all the $187$ suitable candidates. The axes show total magnetization (mag) taken from the Materials Project (MP) database along with the space group (SG), ionic character (ionic char), covalent range (cov range) as calculated from elemental properties, number of elements (num elements) and energy gap (Eg) extracted from MP.} 
    \label{fig:parallel-coordinates-approaches}
\end{figure}

\begin{figure}%[h] %!tbp
    \centering
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/01-ferrenti-approach-v3.pdf}
    \end{subfigure}%
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/02-augmented-ferrenti-approach-v2.pdf}
    \end{subfigure}
    \begin{subfigure}{0.4\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/03-insightful-approach-v3.pdf}
    \end{subfigure}
    %\vspace*{-95mm}
    \caption{Two-dimensional scatter plots for the three different approaches. We have identified the two eigenvectors corresponding to the two largest eigenvalues of the covariance-matrix, that is, the two most important principal components of the initial data from the Materials Project query. Then, we have transformed the three training sets resulting from the three approaches and visualized them as scatter plots. Green squares display suitable candidates, along with the black (SiC), blue (diamond) and yellow (Si) symbols in the bottom panel, and red triangles represent unsuitable candidates.}
    \label{fig:2dscatterplotpca}
\end{figure}

@@

\subsection*{Important features for QT}

@@

%Turning to why the specific materials were chosen ... 
The analysis of the principal components extracted from the machine learning models revealed that eigenvector number five was dominant for both the Ferrenti and extended Ferrenti approaches. This principal component includes features related to the band gap and chemical environment, thus, the band gap criterion imposed in the training set selection is at least somewhat satisfied. On the other hand, the band gap related features were not important in the dominant principal component in the empirical approach, i.e., principal component number one.  
Figure~\ref{fig:histogram_new}(a) contains histograms of the number of predicted suitable materials for all approaches as a function of  band gap (as listen in the MP database), and reveals that all approaches predicted a substantial amount of materials with a low band gap as suitable. Moreover, the distribution of materials that was identified as suitable is rather broad with the Ferrenti based approaches exhibiting a peak around $2.5$~eV. 
Coincidentally, the band gap of, e.g., 4$H$-SiC is usually computed in the range around $2.5$~eV at the PBE level. 
The empirical approach exhibits a more scattered data distribution with local maxima around $1$-$2$ and $4$~eV.

@@

We interpret the standard deviation (std) of the RDF such that zero std in the RDF means that there is no variation in the radial symmetry throughout the material. Similarly, zero std in the average bond length would mean that all bonds are identical throughout the crystal. Intriguingly, the peaks in the distributions of the predicted materials are not found for perfectly symmetric materials with identical bond lengths; instead, some variation in the bond and wavefunction distributions is found to be optimal for a material to be suitable for QT.  
%This is in keeping with no elemental semiconductors being present in the top lists of predicted materials above the $0.75$ confidence level. 
Note that the maxima  in the histograms in Fig.~\ref{fig:histogram_new} seem to appear for moderate standard deviations in the RDF and average bond length, indicating that a certain degree of symmetry is likely necessary for a material to act as a good QT host. The exact degree and type of symmetry is still open to debate and should be studied further.

Interestingly, none of the nine materials predicted by the empirical approach with $0.85$ confidence, nor of the six materials predicted by all approaches to $0.75$ confidence, are elemental. At most three different elements are present, but the emphasis is clearly on binary compounds. This resonates strongly with the observations from the principal component analysis: an optimal degree of crystalline order likely exists for a material to manifest QT compatible properties, but retaining some variations throughout the crystal in (e.g.) bond length and symmetry.  Indeed, this is 
%in contrast 
not precisely as expected according 
to the expectations that guided the formulation of the test, training and validation sets in the Ferrenti and extended Ferrenti approaches. Where most previous works have highlighted features such as band gap, polarity and ionic character as vital for a semiconductor to be a good quantum host, our results instead reveal that local variations in the crystal structure related to symmetry and bond angles, length and orientation should be considered 
%as well. 
instead.

@@

Additionally, the standard deviation of the average bond length (see Figure~\ref{fig:histogram_new}d), site fingerprint of the chemical environment and bond orientation (see the Supplementary Material) are influential in guiding the machine learning algorithms upon classifying suitable and unsuitable materials for QT in the empirical approach.

Considering our findings in light of the specific defect centers we know to be quantum compatible reveals fascinating characteristics. None of the known quantum emitters or spin centers seem to appear in completely uniform systems. For example, in the symmetric materials of diamond and silicon, QT compatible defects are not intrinsic; neither the silicon vacancy in silicon nor the carbon vacancy in diamond, for example,  have exhibited single-photon emission or controllable spin coherence. Instead, we find that QT effects often appear after impurities are introduced, as evidenced by the phosphorous and carbon impurities in silicon [REF] and the nitrogen-vacancy, germanium-vacancy, tin-vacancy an lead-vacancy centers in diamond [REF]. 
For the binary system of silicon carbide in different polytypes, on the other hand, intrinsic defects like the silicon vacancy and the divacancy are appropriate for our goals. 
While these trends have yet to be verified on a grander scale, our findings provide strong indications that local variations in the crystal structure are paramount for QT  compatible properties to manifest in a semiconductor host.

\section*{Discussion} 
We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a data set consisting of more than $25.000$ materials. Furthermore, three approaches to data mining for QT applications were developed; (i) the Ferrenti approach, (ii) the extended Ferrenti approach, and (iii) the empirical approach.  
Three distinct sets of labeled data were defined for training, testing and validation of four different machine learning algorithms (logistic regression, decision trees, random forests and gradient boosting).

We find that the formulation of the training sets for the Ferrenti and extended Ferrenti approaches was broad, and hence multiple dimensions would be required for distinguishing between materials labeled as suitable and unsuitable for QT in the two training sets. The stringent restrictions set for the empirical approach, on the other hand, proved to give a distinct difference between suitable and unsuitable candidates by visualizing the two main principal components of the covariance matrix. 
Moreover, the smaller number of materials predicted for the empirical approach permits manual verification.

Discuss the new findings from the ML here, related to features. Perhaps move down from above.

%NB 
Discuss suitability of ML methods here.

In short, the findings presented herein firmly establish that material informatics (from data mining via featurization to machine learning predictions) is a viable and important route to new discoveries in important fields.  Herein, we focus on predicting new candidate materials to host single-photon emitters and spin centers for quantum technology applications, but the framework we have developed is suitable for other fields as well. 
Two main routes can be followed to exploit the results. One aspect is the pursuit of experimental verification of QT compatible effect in the materials predicted as suitable by the four machine learning methods. Another, perhaps even more important, is to use the features and trends identified during the data mining and prediction process to understand the distinct material characteristics that enable quantum effects to manifest. Thus can new discoveries in the field of quantum technologies be made.

>>>>

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[superscriptaddress,unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
 prb,
 twocolumn,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}

@@

We find that the Ferrenti approach and the extended Ferrenti approach predict a broad range of materials as suitable for quantum technologies. The restrictions set in the empirical approach, which relies exclusively on findings from the literature, gave substantially fewer candidates with a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. 
In contrast to the machine learning (ML) methods based on the Ferrenti and extended Ferrenti approaches that focused on band gap and ionic character, the ML methods from the empirical approach highlighted features related to symmetry and crystal structure, including bond length, orientation and radial distribution, as influential when labeling a material as suitable for QT.  
All four machine learning (ML) methods applied to the training set derived using the empirical approach agreed on a set of $214$ predicted candidate materials.
All three approaches and all four machine learning methods agreed on a subset of $47$ eligible candidates %(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for further material explorations towards quantum technology.

@@

%Figure 1 introduction schematic 
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/ht-workflow-new-2.png}
    \caption{Schematic of an example workflow in material informatics. Results from theory, experiment and computation are fed into material databases (arrow pointing to the center). A cycle involving material screening, machine learning and predictions leads to knowledge gain and ultimately applications in fields such as clean energy and quantum technology. 
    }
    \label{fig:ht-workflow}
\end{figure*}

@@

Here we provide a framework for the data mining and automated discovery of promising solid-state semiconductor hosts for quantum emitters and spin centers using targeted database search and machine learning methods combined with knowledge from the field. 
%\textit{En setning om discovery her??} 
Analyzing the output of the machine learning methods reveals that, given a suitable initial data set of labeled materials for training, testing and validation, it is possible to discern the physical mechanisms that govern a material's suitability for quantum applications.  
In this framework we do not distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that high quality databases containing relevant theoretical and/or experimental data are available.

The approach we have developed relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
The Matminer Python library for data mining \cite{Ward2018} was then used for material analysis to featurize the extracted data. An important aspect of our work is the database building and pertinent development of the training sets for machine learning methods. We have developed three different approaches to data mining: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) 
%Marianne I have now changed augmented Ferrenti to extended Ferrenti 
\emph{the extended Ferrenti approach} and (iii) \emph{the empirical approach}. We find that the former two methodologies base their data extraction protocols on broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The third and empirical approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates.  The three training sets where then analyzed with the four supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting, and yielded $214$ predicted candidates that were common between all approaches and machine learning methods. Importantly, the empirical approach emphasizes features like the standard deviations  of  the  radial  distribution  function  (RDF) and average bond length as critical features for suitable QT compatible materials.
%(to a $0.5$ cut-off)  
Example materials include ZnGeP$_2$, MgSe, CdS, BP, BC$_2$N, BP, Ge, GeC, InP, and InAs. 
Focused theoretical and experimental studies, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, are needed to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart_2.png}
    \caption{The data flow of the main project, starting from an initial MP query, and ending with a featurized data set with entries from several other databases. To limit the memory and computational usage, the data are partitioned into smaller subsets where the respective Pymatgen objects  
    (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) 
    are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. Abbreviations used are Materials Project (MP), density of states (DOS) and electronic band structure (band).
    }
    \label{fig:flowchart}
\end{figure*}

@@

\begin{figure}%[t] %!htp
    \centering
    \begin{subfigure}{0.9\columnwidth}
        \centering
        \input{figures/parallel-coordinates-plots/01-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.9\columnwidth}
        \centering
          \input{figures/parallel-coordinates-plots/02-augmented-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.9\columnwidth}
        \centering
          \input{figures/parallel-coordinates-plots/03-insightful-approach-v2.pgf}
    \end{subfigure}
    %\vspace*{10mm}
    \caption{Parallel coordinate plots for the different approaches. To limit the data cluttering, we have randomly collected up to $250$ entries for each class and made the lines transparent. For the empirical approach, we have used all the $187$ suitable candidates. The axes show total magnetization (mag) taken from the Materials Project (MP) database along with the space group (SG), ionic character (ionic char), covalent range (cov range) as calculated from elemental properties, number of elements (num elements) and energy gap (Eg) extracted from MP.} 
    \label{fig:parallel-coordinates-approaches}
\end{figure}

\begin{figure}%[h] %!tbp
    \centering
    \begin{subfigure}{0.45\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{figures/pca-2d-plots/01-ferrenti-approach-v3.pdf}
    \end{subfigure}%
    \begin{subfigure}{0.45\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{figures/pca-2d-plots/02-augmented-ferrenti-approach-v2.pdf}
    \end{subfigure}
    \begin{subfigure}{0.9\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{figures/pca-2d-plots/03-insightful-approach-v3.pdf}
    \end{subfigure}
    %\vspace*{-95mm}
    \caption{Two-dimensional scatter plots for the three different approaches. We have identified the two eigenvectors corresponding to the two largest eigenvalues of the covariance-matrix, that is, the two most important principal components of the initial data from the Materials Project query. Then, we have transformed the three training sets resulting from the three approaches and visualized them as scatter plots. Green squares display suitable candidates, along with the black (SiC), blue (diamond) and yellow (Si) symbols in the bottom panel, and red triangles represent unsuitable candidates.}
    \label{fig:2dscatterplotpca}
\end{figure}

@@

\section*{Discussion}
%\subsection*{Important features for QT}

@@

%Turning to why the specific materials were chosen ... 
Taking a closer look at the reasoning behind the choices made by the different ML methods during the classification process, we find important differences between the three labeling approaches, and can start to discern the driving forces for manifestation of quantum compatible properties in semiconductors. 
The analysis of the principal components extracted from the machine learning models revealed that eigenvector number five was dominant for both the Ferrenti and extended Ferrenti approaches. This principal component includes features related to the band gap and chemical environment, thus, the band gap criterion imposed in the training set selection is at least somewhat satisfied. 
In other words, the Ferrenti approaches appear to reproduce the logic of the initial selection process, thus, little new information is gained. 
On the other hand, the band gap related features were not important in the dominant principal component in the empirical approach, i.e., principal component number one.  
Figure~\ref{fig:histogram_new}(a) contains histograms of the number of predicted suitable materials for all approaches as a function of  band gap (as listen in the MP database), and reveals that all approaches predicted a substantial amount of materials with a low band gap as suitable. Moreover, the distribution of materials that was identified as suitable is rather broad with the Ferrenti based approaches exhibiting a peak around $2.5$~eV. 
Coincidentally, the band gap of, e.g., 4$H$-SiC is usually computed in the range around $2.5$~eV at the PBE level. 
The empirical approach exhibits a more scattered data distribution with local maxima around $1$-$2$ and $4$~eV.

@@

We interpret the standard deviation (std) of the RDF such that zero std in the RDF means that there is no variation in the radial symmetry throughout the material. Similarly, zero std in the average bond length would mean that all bonds are identical throughout the crystal. Intriguingly, the peaks in the distributions of the predicted materials are not found for perfectly symmetric materials with identical bond lengths; instead, some variation in the bond and wavefunction distributions is found to be optimal for a material to be suitable for QT.  
%This is in keeping with no elemental semiconductors being present in the top lists of predicted materials above the $0.75$ confidence level. 
Note that the maxima  in the histograms in Fig.~\ref{fig:histogram_new} seem to appear for moderate standard deviations in the RDF and average bond length, indicating that a certain degree of symmetry is likely necessary for a material to act as a good QT host. The exact degree and type of symmetry is still open to debate and should be studied further. 
Similarly symmetry related features such as the standard deviation of the average bond length (see Figure~\ref{fig:histogram_new}d), site fingerprint of the chemical environment and bond orientation (see the Supplementary Material) are also influential in guiding the machine learning algorithms upon classifying suitable and unsuitable materials for QT in the empirical approach.

Interestingly, none of the nine materials predicted by the empirical approach with $0.85$ confidence, nor of the six materials predicted by all approaches to $0.75$ confidence, are elemental. At most three different elements are present, but the emphasis is clearly on binary compounds. This resonates strongly with the observations from the principal component analysis: an optimal degree of crystalline order likely exists for a material to manifest QT compatible properties, but retaining some variations throughout the crystal in (e.g.) bond length and symmetry.  Indeed, this is 
%in contrast 
contrary to the expectations that guided the formulation of the test, training and validation sets in the Ferrenti and extended Ferrenti approaches. Where most previous works have highlighted features such as band gap, polarity and ionic character as vital for a semiconductor to be a good quantum host, our results instead reveal that local variations in the crystal structure related to symmetry and bond angles, length and orientation should be considered 
%as well. 
instead.

@@

Considering the trends in machine learning selectivity in light of the specific defect centers we know to be quantum compatible reveals fascinating characteristics. None of the known quantum emitters or spin centers seem to appear in completely uniform systems. For example, in the symmetric materials of diamond and silicon, QT compatible defects are not intrinsic; neither the silicon vacancy in silicon nor the carbon vacancy in diamond, for example,  have exhibited single-photon emission or controllable spin coherence. Instead, we find that QT effects often appear after impurities are introduced, as evidenced by the phosphorous and carbon impurities in silicon [REF] and the nitrogen-vacancy, germanium-vacancy, tin-vacancy an lead-vacancy centers in diamond [REF]. 
For the binary system of silicon carbide in different polytypes, on the other hand, intrinsic defects like the silicon vacancy and the divacancy are appropriate for our goals. 
While these trends have yet to be verified on a grander scale, our findings provide strong indications that local variations in the crystal structure are paramount for QT  compatible properties to manifest in a semiconductor host.

%\section*{Discussion} 
%MB I removed the initial discussion here, as it was a bit repetitive 
%We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a data set consisting of more than $25.000$ materials. Furthermore, three approaches to data mining for QT applications were developed; (i) the Ferrenti approach, (ii) the extended Ferrenti approach, and (iii) the empirical approach. Three distinct sets of labeled data were defined for training, testing and validation of four different machine learning algorithms (logistic regression, decision trees, random forests and gradient boosting).

%NB 
%Morten, kan ta en titt her? 
Discuss suitability of ML methods here.

To summarize, the data extraction tools developed herein resulted in a data set of over $25.000$ materials and more than $4800$ physics-informed features. Three different approaches (Ferrenti, extended Ferrenti and empirical) for data labeling were employed for training, testing and validation of four machine learning algorithms. 
We find a much clearer separation between suitable and unsuitable candidates after data labeling in the empirical approach than for the two Ferrenti based ones. 
Additionally, much fewer principal components are needed for obtaining optimal performance of the machine learning algorithms for the empirical approach, indicating a simpler model. 
The principal component analyses of the ML model's performance imply that the Ferrenti and extended Ferrenti approaches, with their strong focus on band gap and bonding character when assessing a material's quantum compatibility, largely replicate the specifications of the data labeling process. 
Valuable insight is gained from the empirical approach which highlights the importance of symmetry related properties in the bond orientations and wavefunctions instead without being guided by assumptions. 
%that the formulation of the training sets for the Ferrenti and extended Ferrenti approaches was broad, and hence multiple dimensions would be required for distinguishing between materials labeled as suitable and unsuitable for QT in the two training sets. 
%The stringent restrictions set for the empirical approach, on the other hand, proved to give a distinct difference between suitable and unsuitable candidates by visualizing the two main principal components of the covariance matrix. Moreover, the smaller number of materials predicted for the empirical approach permits manual verification.

The findings presented herein firmly establish that material informatics (from data mining via featurization to machine learning predictions) is a viable and important route to new discoveries in important fields.  Herein, we focus on predicting new candidate materials to host single-photon emitters and spin centers for quantum technology applications, but the framework we have developed is suitable for other fields as well. 
Two main routes can be followed to exploit the results. One aspect is the pursuit of experimental verification of QT compatible effect in the materials predicted as suitable by the four machine learning methods. Another, perhaps even more important, is to use the features and trends identified during the data mining and prediction process to understand the distinct material characteristics that enable quantum effects to manifest. Thus can new discoveries in the field of quantum technologies be made.
